Which of my Transient Type Checks are not (Almost) Free?

### 2019-09-11 paper.tex
Added paragraph heading to enable reviewer 2 to find the refined protocol
Signed-off-by: Stefan Marr <git@stefan-marr.de>
Trying to be a bit more correct
Signed-off-by: Stefan Marr <git@stefan-marr.de>

Unlike prior work, we wish to identify which type annotations cause anomalies,
% To investigate Moth's support for transient type checks further, 
and thus we adapted the Takikawa protocol
and took inspiration from the Takikawa-Vitousek variant.
For each benchmark, we generated 100 partially typed versions,
or fewer if the benchmark has less than 11 types.
We did an even split so that for each $i \ge 1$ and $i$ < $N$, we generated roughly the same number of configurations with $i$ type annotations. We used Robert Floyd's sampling algorithm \cite{Bentley:1987:PPS:30401.315746} to randomly choose the type annotations each configuration contained, and we ensured that no duplicate configurations were generated. In addition to these, we tested fully untyped and typed versions, for a total of 102 configurations per benchmark (or 97 in the case of our Storage benchmark, since it only has 10 type annotations).

@@

Of particular note is that some of the graphs (Permute, Storage, Towers, and Richards) show bimodal performance profiles, that is, two separate roughly-horizontal lines. Presumably Moth can remove all the overhead from some configurations, but in others there must be one or more type checks that cannot be optimised away. List shows three performance modes: the graph consists of three mostly flat lines, at about 1 times slower, 1.5 times slower, and 1.8 times slower than untyped code.

@@

In this paper we have investigated how benchmarks can be
repurposed to determine precisely which transient type checks are
likely resistant to the optimisations provided by a just-in-time compiling
virtual machine.
However, detailed analysis will be needed in order to identify exactly what causes such performance effects, such as that we previously did for the List benchmark \cite{roberts-and-co-ecoop-2019}.
We observed that many of our benchmark results
conducted under the Takikawa protocol 
showed a characteristic bimodal performance profile: some
of the benchmark configurations ran significantly slower than
the remaining configurations. We also observed trimodal profiles, as well as performance increases when type annotations where added.

>>>>

\paragraph{Refined Takikawa Protocol.}
Unlike prior work, we wish to identify which type annotations cause anomalies,
% To investigate Moth's support for transient type checks further, 
and thus we adapted the Takikawa protocol
and took inspiration from the Takikawa-Vitousek variant.
For each benchmark, we generated 100 partially typed versions,
or fewer if the benchmark has less than 11 types.
We did an even split so that for each $i \ge 1$ and $i$ < $N$, we generated roughly the same number of configurations with $i$ type annotations. We used Robert Floyd's sampling algorithm \cite{Bentley:1987:PPS:30401.315746} to randomly choose the type annotations each configuration contained, and we ensured that no duplicate configurations were generated. In addition to these, we tested fully untyped and typed versions, for a total of 102 configurations per benchmark (or 97 in the case of our Storage benchmark, since it only has 10 type annotations).

@@

Of particular note is that some of the graphs (Permute, Storage, Towers, and Richards) show bimodal performance profiles, that is, two separate roughly-horizontal lines. Presumably Moth can remove all the overhead from some configurations, but in others there must be one or more type checks that could not be optimised away. List shows three performance modes: the graph consists of three mostly flat lines, at about 1 times slower, 1.5 times slower, and 1.8 times slower than untyped code.

@@

In this paper we have investigated how benchmarks can be
repurposed to determine precisely which transient type checks are
likely to cause performance effects,
and for instance are currently not optimised by the just-in-time compiler.
However, detailed analysis will be needed in order to identify exactly what causes such performance effects, such as that we previously did for the List benchmark \cite{roberts-and-co-ecoop-2019}.
We observed that many of our benchmark results
conducted under the Takikawa protocol 
showed a characteristic bimodal performance profile: some
of the benchmark configurations ran significantly slower than
the remaining configurations. We also observed trimodal profiles, as well as performance increases when type annotations where added.
